---
title: "LogisticRegression - Gradient Descent"
author: "Wal McConnell"
date: "9/29/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logistic Regression

A classic supervised statistical technique for predictive modelling which is based on the linear hypothesis: y = mx + c


In order to apply the linear hypothesis to a dataset with the end aim of modelling the situation under  investigation, there needs to be a linear relationship between the variables in question.


```{r cars}
attach(mtcars)
plot(disp, mpg, col = "blue", pch = 20)
```

Model : Calculate the coefficients by reducing the value of the MSE : Mean Squared Error

```{r}
model <- lm(mpg ~ disp, data = mtcars)
coef(model)
y_preds <- predict(model)
plot(disp, mpg, col = "blue", pch = 20)
abline(model)
```


# Gradient Descent

The Gradient Descent Algorithm which can be used to find the optimal intercept and gradient for any set of data in which a linear relationship exists

```{r}



costs <- data.frame()

gradientDesc <- function(x, y, learn_rate, conv_threshold, n, max_iter) {
  plot(x, y, col = "blue", pch = 20)
  m <- runif(1, 0, 1)
  c <- runif(1, 0, 1)
  
  yhat <- m * x + c
  MSE <- sum((y - yhat) ^ 2) / n
  
  converged = F
  iterations = 0
  
  while(converged == F) {
    ## Implement the gradient descent algorithm
    m <- m - learn_rate * ((1 / n) * (sum((yhat - y) * x)))
    c <- c - learn_rate * ((1 / n) * (sum(yhat - y)))
    yhat <- m * x + c
    MSE_new <- sum((y - yhat) ^ 2) / n
    
    costs <- rbind(costs, data.frame(c(iterations,MSE_new, m, c)))
    
    if(MSE - MSE_new <= conv_threshold) {
      abline(c, m) 
      converged = T
      return(paste("Optimal intercept:", c, "Optimal slope:", m))
    }
    iterations = iterations + 1
    if(iterations > max_iter) { 
      abline(c, m) 
      converged = T
      return(paste("Optimal intercept:", c, "Optimal slope:", m))
    }
  }
}


#gradientDesc(disp, mpg, 0.0000293, 0.001, 32, 2500000)
gradientDesc(disp, mpg, 0.0000293, 0.001, 32, 250)

nrow(costs)

```

