---
title: "Linear Regression & Gradient Descent"
author: "Wal McConnell"
date: "9/29/2017"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(reshape2)
knitr::opts_chunk$set(echo = TRUE)
```

A supervised statistical technique for predictive modelling which is based on the linear hypothesis: y = mx + c. Therefore there needs to be a linear relationship between the variables in question.


```{r cars}
x <- runif(1000, 0, 10)
y <- x + rnorm(1000) + 3
data <- data.frame(x=x, y=y)
```

# Simple Linear Model

``` {r}
model <- lm(y ~ x, data)
model
ggplot(data, aes(x, y)) + 
  geom_point(color="grey74", shape=21) +
  geom_line(aes(y=fitted(model)), color = "blue")
```


# Gradient Descent

The Gradient Descent Algorithm finds the optimal intercept and gradient for any set of data in which a linear relationship exists

```{r}

# squared error cost function
cost <- function(X, y, theta) {
  sum( (X %*% theta - y)^2 ) / (2*length(y))
}

# learning rate and iteration limit
alpha <- 0.01
num_iters <- 1000

# keep history
cost_history <- double(num_iters)
theta_history <- list(num_iters)

# initialize coefficients
theta <- matrix(c(0,0), nrow=2)

# add a column of 1's for the intercept coefficient
X <- cbind(1, matrix(x))

# gradient descent
for (i in 1:num_iters) {
  error <- (X %*% theta - y)
  delta <- t(X) %*% error / length(y)
  theta <- theta - alpha * delta
  
  cost_history[i] <- cost(X, y, theta)
  theta_history[[i]] <- theta
}

print(theta)
```



```{r}
plot <- ggplot(data, aes(x, y)) + geom_point(color="grey")

for (i in c(1,2,3,5,7,seq(10,num_iters,by=10))) {
  plot <- plot + 
    geom_abline(intercept=theta_history[[i]][1], 
                slope=theta_history[[i]][2], 
                color="brown2", alpha=0.2)
}

plot
```


## Cost Function
```{r}
cost_data <- melt(cost_history)

ggplot(cost_data, aes(x = 1:nrow(cost_data), y = value)) + 
  geom_line() +
  labs(x = "Iterations", y = "Cost") +
  ggtitle("Cost Function Reduction")
```

# Prediction

```{r}
px1 <- 4.21
py1 <- theta[1] + (px1*theta[2])

px2 <- 9.01
py2 <- theta[1] + (px2*theta[2])


predictions <- data.frame(x=c(px1,px2), y=c(py1,py2))

ggplot(data, aes(x, y)) + 
  geom_point(color="grey74", shape=21) +
  geom_line(aes(y=fitted(model)), color = "blue") +
  geom_point(data=predictions, aes(x=x, y=y), color="red", size=5, shape=4)

```

